---
title: "Challenge 3"
author: "Eva Dykaar"
format: 
  html:
    self-contained: true
    code-tools: true
    code-fold: true
execute:
  echo: true
  error: true
  message: false
---

# Challenge 3: Group Comparisons & Data Ethics

```{r packages, message = FALSE}
# code chunk for loading packages and importing the data
#| message: false

library(tidyverse)
library(here)
hiphop <- read_csv(here("Week 3", "hiphop.csv"))

hiphop_clean <- hiphop |>
  mutate(
    across(
      c(word, sex, ethnic, familiarity), as.factor
    )
  )

hiphop_clean <- hiphop_clean|>
  mutate(ethnic = case_when(ethnic == "white" ~ "White", TRUE ~ "Non-White"))

```


## Published Comparisons

In the published article ([You Know What It Is: Learning Words through Listening to Hip-Hop](https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0028248&type=printable)), the author presents a series of comparisons about the variables which most explain an individual's familiarity with African American English (AAE).

Let us instead compare the number of artists participants reported liking for each genre. Specifically, you will determine which music genre most differentiates each of the demographic groups provided.

> Which genre had much higher average (mean or median) reported artists in one group than the other.
>
> -   Male versus Female
> -   White versus Non-White

::: callout-tip
You might find it helpful to first create a new data set with only the variables you are interested in! Look at the Music Variables in the data set description along with the demographics of interest.

Helpful functions: `select()`, `group_by()`, `summarize()`, `across()`

Other useful operations in R: `mean()`, `diff()`, `abs()`, `which.max()`
:::

```{r}
newhiphop <- hiphop_clean |>
  select(c(sex, ethnic, intl:unclassifiable))

newhiphop_sex <- newhiphop |>
  group_by(sex) |>
  summarise(
    across(
      intl:unclassifiable, 
      mean, 
      na.rm = TRUE
      )
    )
    
newhiphop_sex <- newhiphop_sex |>
  summarise(
    abs(
    across(
      intl:unclassifiable,
      diff, na.rm = TRUE
    )
  )
  )
newhiphop_sex |>
  which.max()
```
The genre rock has a higher average reported artists in males vs females.


```{r}
newhiphop_ethnic <- newhiphop |>
  group_by(ethnic) |>
  summarise(
    across(
      intl:unclassifiable, 
      mean, 
      na.rm = TRUE
      )
    )

newhiphop_ethnic <- newhiphop_ethnic |>
  summarise(
    abs(
    across(
      intl:unclassifiable,
      diff, na.rm = TRUE
    )
  )
  )

newhiphop_ethnic |>
  which.max()
```
The genre international music has a higher average reported artists in white vs non-white.


## Study Design Critique -- Data Ethics

Myself, members of the Tidy Tuesday community, and previous 331 students have voiced concerns regarding the design and social context of this study.

You've already read the [data description (link)](http://conservancy.umn.edu/bitstream/handle/11299/116327/5/explanationAAEHiphopChesley.txt) regarding how participants were recruited for participation in this study. **Now**, you need to read additional details regarding aspects of the study in the published paper: [You Know What It Is: Learning Words through Listening to Hip-Hop](https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0028248&type=printable).

> Based on the design of this study and its context (African American English), what are **at least two** concerns you have? Keep in mind this critique aligns with conversations regarding data ethics. Thus, your concerns need to address the racial aspects of the design of this study.

The first concern about the racial aspects of the study is extrapolating the data to a larger population. This data is from a small sample size and from one university, so I would be careful when generalizing the results because they might not be accurate with a bigger population. Another concern about the racial aspects of the design of this study is there could be many more confounding variables besides just the music choice of each participant and to not include many other variables is concerning. More concerns about this study are the implications of stereotyping AAE and how the data is based on popular music and its trend in society and how the participants are supposed to be getting their definitions from popular media sources. Also, the people that were recruited into the study were all not African American.

Re-submission Reflection:
My thinking changed because I realized that I did not explain the root caused of the racial concerns of the study, stereotyping, and I needed to explain how this is ethically concerning. I also needed to think about who was recruited to this study and what was reported.



# Challenge 3 Submission

Your challenge should be submitted as a **separate file**, **not** at the bottom of your Lab 3 file. Please submit your rendered HTML file. You can copy and paste this code into a **new** Quarto file. Your Challenge 3 submission should only included code necessary for completing the Challenge, nothing else.

You will submit **only** your rendered HTML file. Your HTML file is required to have the following specifications in the YAML options (at the top of your document):

-   be self-contained (`self-contained: true`)
-   include your source code (`code-tools: true`)
-   include all your code and output (`echo: true`)
-   include **no** messages from loading in packages or the data (`messages: false`)

**If any of the options are not included, your Lab 3 or Challenge 3 assignment will receive an "Incomplete" and you will be required to submit a revision.**
